SDR enhancement by clouds

### 2024-03-26 article/article.Rmd
Commit 2024-03-24 21:12
Commit 2024-03-26 11:10

@@

To be able to detect the CE cases, we had to establish a baseline, above which we can
characterize each data point as an enhancement event, by estimating the occurring
over irradiance (OI).  To have an estimation and a first insight on the phenomenon,
we experimented with two simple approaches for the reference.  The Haurwitz's model
[@Haurwitz1945], which is as simple clear sky model, and we had already adjusted and
had good fit with our data [@Natsis2023], and the TSI at the top of the atmosphere.
We have tested both cases by using an appropriate relative threshold and/or an
additional constant offset.  The initial results, showed that we can detect a big
portion of the CE events. These results were helpful, and helped to establish some
criteria to further improve the CE identification. It was evident, by inspecting the
daily plot of irradiance, that changes on the atmospheric conditions introduced
numerous false positive and false negative results.  To produce a more accurate
reference, we had to take into account more factors that effect the clear sky
radiation. So we used a radiation transfer model.

@@

<!-- reasoning for cretria -->
We have to note, that the differentiation of the threshold factor was needed, because
of the high irradiance values we observed, early in the morning and late on the
afternoon. We have confirmed, by inspecting images from the sky cam, that we have
duration of elevated irradiance, either due the clearness of the atmosphere, or some
interferences by reflections on nearby bright surfaces. Although, these cases may
mask some of the CE events. The actual CE events, produce higher irradiance and thus
are identified as such. As a side effect, the reported OI will be slight
underestimated for those SZAs, but the overall contribution of those cases to the
total daily energy, is minimal due the low occurrences and the lower irradiances.

@@

These selections has some subjectivity, as the definition of clear sky, is depended
on the intended usage. Our main focus is to be able to identify the OI created by the
clouds. Although, the aerosols and water vapor are always present and can not be
completely removed. 
The effect of the different clearness levels

@@

Our dataset, after the data selection processing, consists of
`r ST_total$wattGLB.N` records of GHI, of which
$`r round(100*ST_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ are CE and
$`r round(100*ST_extreme_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ are ECE events.
The highest GHI recorder was 
$`r round(ST_daily[which.max(wattGLB.max), wattGLB.max],1)`\,W/m^2$
on `r strftime(ST_daily[which.max(wattGLB.max), Date], "%d\\ %B %Y")`.
The absolute stronger CE event had an over irradiance of 
$`r round(ST_daily[which.max(GLB_diff.max), GLB_diff.max], 2)`\,W/m^2$ on
`r strftime(ST_daily[which.max(GLB_diff.max), Date], "%d\\ %B %Y")`.
The relative stronger CE event was 
$`r round(100*ST_daily[which.max(GLB_ench.max), GLB_ench.max],1)`\,\%$ above the 
clear sky threshold, on 
`r strftime(ST_daily[which.max(GLB_ench.max), Date], "%d\\ %B %Y")`.

@@

We computed the daily trend of the mean over irradiance of CE
(Figure\nobreakspace{}\ref{fig:CEmeanDaily}), using a first-order autoregressive
model with lag of 1 day, using the 'maximum likelihood' fitting method [@Gardner1980;
@Jones1980] by implementing the function 'arima' from the library 'stats' of the R
programming language [@RCT2023]. The trends are reported together with the $2\sigma$
error.  We observe an increase of
$`r round(dailyCE[var == "GLB_diff.mean", Tmod_Estimate], 3)`\pm
`r round(2 * dailyCE[var == "GLB_diff.mean", Tmod_Std..Error], 3)`\,W/m^2$
on the mean over irradiance.

@@

Although, the previous result is closer to the raw data, we preferred to present the
annual statistics, that give a more clear picture about the long term CE trends.
Hence, we calculated the annual values from the one-minute measurements.  The annual
mean over irradiance is
$`r round(yearlyCE[var == "GLB_diff.mean", slope], 2)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 2)`\,W/m^2$
(Figure\nobreakspace{}\ref{fig:P-energy-mean}),
but this value alone is not very useful due the intrinsic high variability of this
metric.

@@

```{r P-energy-mean, echo=FALSE, fig.cap="Trends of the mean over irradiance per CE."}
knitr::include_graphics("../images/P_energy-7.png")
```

@@

In order to further study the characteristics of the CE events, we grouped the singe
minute CE events, to continuous CE groups (CEG). Thus, a CEG consists of one or more
successive CE cases, and can have a duration of multiple of minutes.  We have
identified `r ST_G0[, .N]` groups, where the group with the longest duration lasted
`r ST_G0[, max(GLB_diff.N)]` minutes on 
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can conclude that longer durations are getting increasingly rare, with durations above 10 minutes very rare.

>>>>

The shortwave solar irradiance, reaches the Earth's surface, is the main energy
source of the planet. It drives and governs the climate.
It has direct practical application, in different energy production method.  The
variability of its intensity, can cause difficulties in predicting the yield and
designing the appropriate equipment. An aspect of this variability are the
interaction

@@

To be able to detect the CE cases, we had to establish a baseline, above which we can
characterize each data point as an enhancement event, by estimating the occurring
over irradiance (OIR).  To have an estimation and a first insight on the phenomenon,
we experimented with two simple approaches for the reference.  The Haurwitz's model
[@Haurwitz1945], which is as simple clear sky model, and we had already adjusted and
had good fit with our data [@Natsis2023], and the TSI at the top of the atmosphere.
We have tested both cases by using an appropriate relative threshold and/or an
additional constant offset.  The initial results, showed that we can detect a big
portion of the CE events. These results were helpful, and helped to establish some
criteria to further improve the CE identification. It was evident, by inspecting the
daily plot of irradiance, that changes on the atmospheric conditions introduced
numerous false positive and false negative results.  To produce a more accurate
reference, we had to take into account more factors that effect the clear sky
radiation. So we used a radiation transfer model.

@@

<!-- reasoning for cretria -->
We have to note, that the differentiation of the threshold factor was needed, because
of the high irradiance values we observed, early in the morning and late on the
afternoon. We have confirmed, by inspecting images from the sky cam, that we have
duration of elevated irradiance, either due the clearness of the atmosphere, or some
interferences by reflections on nearby bright surfaces. Although, these cases may
mask some of the CE events. The actual CE events, produce higher irradiance and thus
are identified as such. As a side effect, the reported OIR will be slight
underestimated for those SZAs, but the overall contribution of those cases to the
total daily energy, is minimal due the low occurrences and the lower irradiances.

@@

These selections has some subjectivity, as the definition of clear sky, is depended
on the intended usage. Our main focus is to be able to identify the OIR created by the
clouds. Although, the aerosols and water vapor are always present and can not be
completely removed. 
The effect of the different clearness levels

@@

Our dataset, after the data selection processing, consists of
`r ST_total$wattGLB.N` records of GHI, of which
$`r round(100*ST_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ are CE and
$`r round(100*ST_extreme_total$GLB_diff.N_pos/ST_total$wattGLB.N, 3)`\,\%$ are ECE events.
The highest GHI recorder was 
$`r round(ST_daily[which.max(wattGLB.max), wattGLB.max],1)`\,W/m^2$
on `r strftime(ST_daily[which.max(wattGLB.max), Date], "%d\\ %B %Y")`.
The absolute stronger CE event had an OIR of 
$`r round(ST_daily[which.max(GLB_diff.max), GLB_diff.max], 2)`\,W/m^2$ on
`r strftime(ST_daily[which.max(GLB_diff.max), Date], "%d\\ %B %Y")`.
The relative stronger CE event was 
$`r round(100*ST_daily[which.max(GLB_ench.max), GLB_ench.max],1)`\,\%$ above the 
clear sky threshold, on 
`r strftime(ST_daily[which.max(GLB_ench.max), Date], "%d\\ %B %Y")`.

@@

We computed the daily trend of the mean OIR of CE
(Figure\nobreakspace{}\ref{fig:CEmeanDaily}), using a first-order autoregressive
model with lag of 1 day, using the 'maximum likelihood' fitting method [@Gardner1980;
@Jones1980] by implementing the function 'arima' from the library 'stats' of the R
programming language [@RCT2023]. The trends are reported together with the $2\sigma$
error.  We observe an increase of
$`r round(dailyCE[var == "GLB_diff.mean", Tmod_Estimate], 3)`\pm
`r round(2 * dailyCE[var == "GLB_diff.mean", Tmod_Std..Error], 3)`\,W/m^2$
on the mean OIR.

@@

Although, the previous result is closer to the raw data, we preferred to present the
annual statistics, that give a more clear picture about the long term CE trends.
Hence, we calculated the annual values from the one-minute measurements.  The annual
mean OIR is
$`r round(yearlyCE[var == "GLB_diff.mean", slope], 2)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.mean", slope.sd], 2)`\,W/m^2$
(Figure\nobreakspace{}\ref{fig:P-energy-mean}),
but this value alone is not very useful due the intrinsic high variability of this
metric.

@@

```{r P-energy-mean, echo=FALSE, fig.cap="Trends of the mean OIR per CE."}
knitr::include_graphics("../images/P_energy-7.png")
```

@@

In order to further study the characteristics of the CE events, we grouped the single
minute CE events, to continuous CE groups (CEG). Thus, a CEG consists of one or more
successive CE cases, and can have a duration of multiple of minutes.  We have
identified `r ST_G0[, .N]` groups, where the group with the longest duration lasted
`r ST_G0[, max(GLB_diff.N)]` minutes on
`r strftime(ST_G0[which.max(GLB_diff.N), Date], "%d\\ %B %Y")`.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can conclude that
longer durations are getting increasingly rare, with durations above 10 minutes very
rare.
